---
title: "README"
author: "kat j"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    preserve_yaml: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **T**ranscript **A**nalysis of **R**NA **Va**riants (TARVa)                                                               
## Project Overview                                   
  [**TARVa**](#transcript-analysis-of-rna-variants-tarva) is a tool which enables analysis of epitranscriptomic modifications. Starting with a global overview, the pipeline moves through non-hierarchical stratifications of the data, ending with local, per-position comparisons across tissue types and conditions. The  workflow is set up for use with paired RNAseq and WGS raw sequencing datasets (fastq) from more than one condition and/or tissue, and covers three data preprocessing pipelines as well as the main TARVa pipeline.                                                         

## Directory Structure                                             

<details>
<summary><strong>📂 Project Directory Tree (click to expand)</strong></summary>

```{r structure-tree, echo=FALSE, results='asis'}
cat("<pre><code>")
cat(readLines("structure.txt"), sep = "\n")
cat("</code></pre>")
```

</details>                                            

## Tools, Environments and Dependencies                                             
For __data preprocessing pipelines__ ([gatk_rna](GatkRnaSeqPipe/), [gatk_wgs](GatkWGSeqPipe/), [stringtie](StringtiePipe/)):  
  - conda envs
    - *bio_qc* ([text](bio_qc_channels_dependencies.txt), [yml](bio_qc.yml))                            
    

## Getting Started                                
After installing the necessary tools and dependencies, 
                          
## How to Use                                   
## Limitations and Considerations                                       
## Acknowledgments                                 
## License                                    




****** NOTES NOTES NOTES*******************************************************************************************************************
TARVa is a tool that identifies, analyzes, and compares RNA-editing events between two conditions.  This approach takes into consideration both depth of read and length of transcript, and scans for "A-I" and "C-U" variants in a strand-aware approach. TARVa also has an optional feature, for studies which provide both RNA-seq and WGS-seq data for each sample, to automatically filter out variant calls that are a result of the DNA transcription process and not of post-modification (RNA-editing) events.                                                              

###### **code draft suggestion 1; create user-input variable functions for paths, sample identifier information, and condition names**                                                                  
###### **code draft suggestion 2: make necessary changes to reflect the optionality of genomic variant calls being added to the analysis pipeline**                            

#### Setup                                                                                                      


See the "About.pdf" document for details on the \underline{data processing and preparation pipeline}. Not following this pipeline might lead to inaccurate outputs from TARVa                     

##### 1.) Unzip TARVaScripts.zip                                                              

&nbsp;&nbsp;&nbsp;&nbsp;Contents:                           
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**a.** *BuildTarvaDBs.py*                           
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; This is the **main script**. *make_sample_tabs*,*each_gene*, *raw_counts*, *dictionaries* and *analyze_lens* are all static methods, housed also in the TARVaScripts directory.                                                                                                                                     
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**b.** *TARVa.slurm*                                  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;This is the slurm directive script that activates the **tarva** environment, assigns path information and runs the main script.                               

**code draft suggestion 3: remove this from scripts directory and instead provide information in a later section of this document describing resources that are needed**                                                        

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**c.** *tarva_env_deps.txt*                                   
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;TARVa runs in the custom conda environment, **tarva**. This file indicates all of the packages and dependencies that belong in the environment. To create the environment, [See step 2](#2-create-the-tarva-environment)                                                                

##### 2.) Create the Tarva environment                            

The **tarva** environment can be created using the command:                                       
`conda create -n [env_name] -c [channel] package(s)` c                          
 
##### 3.) Assign path names and condition names in scripts ----->>>>> REMOVE THIS STEP ONCE code draft suggestion 1 is implemented                                    

##### 4.) Submit job directive script
