---
title: "README"
author: "kat j"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    preserve_yaml: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **T**ranscript **A**nalysis of **R**NA **Va**riants (TARVa)                                                               
## Project Overview                                   
  [**TARVa**](#transcript-analysis-of-rna-variants-tarva) is a tool which enables analysis of epitranscriptomic modifications from raw RNA sequencing data, taking  into consideration both depth of read and length of transcript, and scans for all variant call-types in a strand-aware fashion. Starting with a global overview, the pipeline moves through non-hierarchical stratifications of the data, ending with local, per-position comparisons across tissue types and conditions. The  workflow is set up for use with paired-sample raw RNAseq and whole genome sequencing files (fastq) from more than one condition and/or tissue, and covers three data preprocessing pipelines in addition to the main TARVa analysis pipeline.                                                         

## Tools, Environments and Dependencies
Python 3 is a requirement for this workflow and custom conda environments were created for each group of pipelines ([data pre-processing](#data-pre-processing-pipelines), [TARVa analysis](#tarva-analysis-pipelines)).                                                

For [__data pre-processing pipelines__](#data-pre-processing-pipes)<a name="data-pre-processing-pipelines"></a>  ([Stringtie](StringtiePipe/), [GATK WGS](GatkWGSeqPipe/), [GATK RNA](GatkRnaSeqPipe/)):  
  - conda envs                                                                        
      -- *bio_qc* dependencies ([text](bio_qc_channels_dependencies.txt), [yml](bio_qc.yml))    
  
  - R/4.3.3                                                                     
  
  - samtools/1.11                                        
  
  - gatk/4.2.0                                                    
  
  - STAR                                                  
  
  - Stringtie 
  
  - BWA
  
For the [__TARVa analysis pipelines__](#tarva-analysis-pipes)<a name="tarva-analysis-pipelines"></a> ([DB Build](TARVaCreation/OriginalBuild/), [Analysis set 1](TARVaCreation/DownStreamDBCleanup/),  [Analysis set 2](TARVaCreation/SecondSetBuild/)):                                                           
  - conda envs                                                                            
    -- *tarva* dependencies ([text](tarva_channels_dependencies.txt), [yml](tarva.yml))                                                               
  
  - python modules                                            
    -- *sqlite3*                                                                                                                                          
    -- *numpy*                                       
    -- *scipy*                                       
    -- *pandas*                                    
    -- *statsmodels*                               
    -- *vcf*                                               
                          
## Directory Structure                                             

<details>
<summary><strong>📂 Project Directory Tree (click to expand)</strong></summary>

```{r structure-tree, echo=FALSE, results='asis'}
cat("<pre><code>")
cat(readLines("structure.txt"), sep = "\n")
cat("</code></pre>")
```

</details>                   

## Workflow Overview                                                                                                             
To implement the full TARVa workflow, the following [__data pre-processing pipelines__](#data-pre-processing-pipelines)<a name="data-pre-processing-pipes"></a> are run in the following order:                                                   
  1. [Stringtie pipeline](StringtiePipe/) ([README](StringtiePipe/README_Stringtie.md))                                    
    Stringtie receives STAR RNA sequencing alignment bam files as input and creates transcript expression level files (.ctab) and a gtf file containing nucleotide positions, transcript features, etc. of each transcript identified in the dataset.                                             

  2. [GATK WGS](GatkWGSeqPipe/) ([README](GatkWGSeqPipe/README_gatkWGS.md))                                       
    A GATK pipeline for small somatic variant-finding in WGS sequencing data is implemented to provide the TARVa tool with a baseline set of somatic genomic variants (vcf as output) for each sample. WGS data only for samples which also have RNAseq data available is used in this step.                                                                               

  3. [GATK RNA](GatkRnaSeqPipe/) ([README](GatkRnaSeqPipe/README_gatkRNA.md))                                         
    For samples with WGS data that has been processed in the previous step, RNAseq sequencing data is run through a GATK pipeline for small variant finding in RNAseq (vcf as output).                                                                     
  

Output from the data pre-processing phase of the workflow can then be used as input for the [__TARVa analysis pipelines__](#tarva-analysis-pipelines)<a name="tarva-analysis-pipes"></a>, which can be executed in the following order:                                       

  1. [DB Build](TARVaCreation/OriginalBuild/) ([README](TARVaCreation/OriginalBuild/README_OriginalBuild.md))                         
    RNA variants are first parsed to filter out any genomic variants, then a database is built from several data sources to store detailed information about RNA-specific modifications in the dataset.                                                 
  
  2. [Analysis set 1](TARVaCreation/DownStreamDBCleanup/) ([README](TARVaCreation/DownStreamDBCleanup/README_set1_downstream.md))                         
    Analyses of first set of data, consisting of one tissue type, two different conditions.                                                                        
    
  3. [Analysis set 2](TARVaCreation/SecondSetBuild/) ([README](TARVaCreation/SecondSetBuild/README_secondset_build.md))                                         
    Analyses of second set of data, consisting of a third condition for the tissue type assessed in the previous step and the three same conditions for a different tissue.                                           
    
## Limitations and Considerations                                                   
-     This tool was created for and tested on datasets from the [__Religious Orders Study/Memory and Aging Project (ROSMAP)__](https://dss.niagads.org/cohorts/religious-orders-study-memory-and-aging-project-rosmap/). While the code works for the ROSMAP dataset in computing envrinments comparable to the testing environment,  it is not yet generalizable to other datasets or computing environments. Further more, Analysis Set 2 was added on later in the project but could be integrated with Analysis set 1 so that all conditions and tissues can be assessed simultaneously.                             
-   Current goals for development include the following:                                    
  --    fix code and database redundancies                                                   
  --    replace absolute paths with relative paths                                                
  --    generalizability of condition, tissue, and related variables for applicability to different datasets                                                
  

## License                                    


