---
title: "README"
author: "kat j"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    preserve_yaml: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **T**ranscript **A**nalysis of **R**NA **Va**riants (TARVa)                                                               
## Project Overview                                   
  [**TARVa**](#transcript-analysis-of-rna-variants-tarva) is a tool which enables analysis of epitranscriptomic modifications from raw RNA sequencing data, taking  into consideration both depth of read and length of transcript, and scans for all variant call-types in a strand-aware fashion. Starting with a global overview, the pipeline moves through non-hierarchical stratifications of the data, ending with local, per-position comparisons across tissue types and conditions. The  workflow is set up for use with paired-sample raw RNAseq and whole genome sequencing files (fastq) from more than one condition and/or tissue, and covers three data preprocessing pipelines in addition to the main TARVa analysis pipeline.                                                         

## Tools, Environments and Dependencies
Python 3 is a requirement for this workflow and custom conda environments were created for each group of pipelines ([data preprocessing](#data-preprocessing-pipelines), [TARVa analysis](#tarva-analysis-pipelines)).                                                

For [__data preprocessing pipelines__](#data-preprocessing-pipes)<a name="data-preprocessing-pipelines"></a>  ([Stringtie](StringtiePipe/)<a name="stringtie-pipeline"></a>, [GATK WGS](GatkWGSeqPipe/)<a name="gatk-wgs"></a>, [GATK RNA](GatkRnaSeqPipe/))<a name="gatk-rna"></a>:  
  - conda envs                                                                        
      -- *bio_qc* dependencies ([text](bio_qc_channels_dependencies.txt), [yml](bio_qc.yml))    
  
  - R/4.3.3                                                                     
  
  - samtools/1.11                                        
  
  - gatk/4.2.0                                                    
  
  - STAR                                                  
  
  - Stringtie 
  
  - BWA
  
For the [__TARVa analysis pipelines__](#tarva-analysis-pipes)<a name="tarva-analysis-pipelines"></a> ([Analysis set 1](TARVaCreation/OriginalBuild/), [Analysis set 1 checks and changes](TARVaCreation/DownStreamDBCleanup/),  [Analysis set 2](TARVaCreation/SecondSetBuild/)):                                                           
  - conda envs                                                                            
    -- *tarva* dependencies ([text](tarva_channels_dependencies.txt), [yml](tarva.yml))                                                               
  
  - python modules                                            
    -- *sqlite3*                                                                                                                                          
    -- *numpy*                                       
    -- *scipy*                                       
    -- *pandas*                                    
    -- *statsmodels*                               
    -- *vcf*                                               
                          
## Directory Structure                                             

<details>
<summary><strong>📂 Project Directory Tree (click to expand)</strong></summary>

```{r structure-tree, echo=FALSE, results='asis'}
cat("<pre><code>")
cat(readLines("structure.txt"), sep = "\n")
cat("</code></pre>")
```

</details>                   

## Workflow Overview                                                                                                             
To implement the full TARVa workflow, the following [__data pre-processing pipelines__](#data-preprocessing-pipelines)<a name="data-preprocessing-pipes"></a> are run in the following order:                                                   
  1. [Stringtie pipeline](StringtiePipe/) ([README](StringtiePipe/README_Stringtie.md))                                    
    Stringtie receives STAR RNA sequencing alignment bam files as input and creates transcript expression level files (.ctab) and a gtf file containing nucleotide positions, transcript features, etc. of each transcript identified in the dataset.                                             

  2. [GATK WGS](GatkWGSeqPipe/) ([README](GatkWGSeqPipe/README_gatkWGS.md))                                       
    A GATK pipeline for small somatic variant-finding in WGS sequencing data is implemented to provide the TARVa tool with a baseline set of somatic genomic variants (vcf as output) for each sample. WGS data only for samples which also have RNAseq data available is used in this step.                                                                               

  3. [GATK RNA](GatkRnaSeqPipe/) ([README](GatkRnaSeqPipe/README_gatkRNA.md))                                         
    For samples with WGS data that has been processed in the previous step, RNAseq sequencing data is run through a GATK pipeline for small variant finding in RNAseq (vcf as output).                                                                     
  

Output from the data pre-processing phase of the workflow can then be used as input for the [__TARVa analysis pipelines__](#tarva-analysis-pipelines)<a name="tarva-analysis-pipes"></a>, which can be executed in the following order:                                       

  1. [Analysis set 1](TARVaCreation/OriginalBuild/) ([README](TARVaCreation/OriginalBuild/README_OriginalBuild.md))                         
    RNA variants are first parsed to filter out any genomic variants, then a database is built from several data sources to store detailed information about RNA-specific modifications in the dataset.                                                 
  
  2. [Analysis set 1 checks and changes](TARVaCreation/DownStreamDBCleanup/) ([README](TARVaCreation/DownStreamDBCleanup/README_set1_downstream.md))                                          
  3. [Analysis set 2](TARVaCreation/SecondSetBuild/) ([README](TARVaCreation/SecondSetBuild/README_secondset_build.md))                                         

## Limitations and Considerations                                                   


## License                                    


