---
title: "README"
author: "kat j"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    preserve_yaml: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **T**ranscript **A**nalysis of **R**NA **Va**riants (TARVa)                                                               
## Project Overview                                   
  [**TARVa**](#transcript-analysis-of-rna-variants-tarva) is a tool which enables analysis of epitranscriptomic modifications from raw RNA sequencing data, taking  into consideration both depth of read and length of transcript, and scans for all variant call-types in a strand-aware fashion. Starting with a global overview, the pipeline moves through non-hierarchical stratifications of the data, ending with local, per-position comparisons across tissue types and conditions. The  workflow is set up for use with paired RNAseq and WGS raw sequencing datasets (fastq) from more than one condition and/or tissue, and covers three data preprocessing pipelines as well as the main TARVa pipeline.                                                         

## Directory Structure                                             

<details>
<summary><strong>📂 Project Directory Tree (click to expand)</strong></summary>

```{r structure-tree, echo=FALSE, results='asis'}
cat("<pre><code>")
cat(readLines("structure.txt"), sep = "\n")
cat("</code></pre>")
```

</details>                                            

## Tools, Environments and Dependencies
Python 3 is a requirement for this workflow.                                                  

For __data preprocessing pipelines__ ([gatk_rna](GatkRnaSeqPipe/), [gatk_wgs](GatkWGSeqPipe/), [stringtie](StringtiePipe/)):  
  - conda envs                                                                        
      -- *bio_qc* ([text](bio_qc_channels_dependencies.txt), [yml](bio_qc.yml))    
  
  - R/4.3.3                                                                     
  
  - samtools/1.11                                        
  
  - gatk/4.2.0                                                    
  
  - STAR                                                  
  
  - Stringtie 
  
  - BWA
  
For the __TARVa analysis pipelines__ ([analysis set 1](TARVaCreation/OriginalBuild/), [analysis set 1: checks and changes](TARVaCreation/DownStreamDBCleanup/),  [analysis set 2](TARVaCreation/SecondSetBuild/)):                                                           
  - conda envs                                                                            
    -- *tarva* ([text](tarva_channels_dependencies.txt), [yml](tarva.yml))                                                               
  
  - python modules                                            
    -- *sqlite3*                                                                                                                                          
    -- *numpy*                                       
    -- *scipy*                                       
    -- *pandas*                                    
    -- *statsmodels*                               
    -- *vcf*                                               
                          
## Workflow Overview                                                                                                             
To implement the full TARVa workflow, the __data pre-processing pipelines__ ([gatk_rna](GatkRnaSeqPipe/) ([README](GatkRnaSeqPipe/README_gatkRNA.md)), [gatk_wgs](GatkWGSeqPipe/) ([README](GatkWGSeqPipe/README_gatkWGS.md)), [stringtie](StringtiePipe/) ([README](StringtiePipe/README_Stringtie.md))) are run first.                               
Output from each pipeline can then be used as input for the __TARVa analysis pipeline__ ([analysis set 1](TARVaCreation/OriginalBuild/) ([README](TARVaCreation/OriginalBuild/README_OriginalBuild.md)), [analysis set 1: checks and changes](TARVaCreation/DownStreamDBCleanup/) ([README](TARVaCreation/DownStreamDBCleanup/README_set1_downstream.md)), [analysis set 2](TARVaCreation/SecondSetBuild/) ([README](TARVaCreation/SecondSetBuild/README_secondset_build.md))).                                              


## Limitations and Considerations                                                       

## Acknowledgments                                 
## License                                    


